---
title: "Get started with benchtm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with benchtm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  cache = TRUE
)
```
R package "benchtm" can be used to generate data for the comparison of subgroup identification algorithms. To start with, we can directly load the library using:

```{r}
library(benchtm)
library(dplyr)
```


## Data Generation Framework
Consider a two-arm study where trt = 0 represents control arm and trt = 1 represents treatment arm, the clinical outcome/response $Y$ is generated from 

\begin{equation}
\mathbb{E} (Y|X) = g(f(X)) = g(f_{prog}(X) + trt*(\beta_0 + \beta_1 f_{pred}(X))),
\end{equation}

where $g(\cdot)$ is the link function and $f_{prog}(X)$ is the functional form of the prognostic variables and $f_{pred}(X)$ is the functional form of the predictive variables; $\beta_0$ and $\beta_1$ are the coefficients of main treatment effect and the predictive effects respectively. From the data generation formula one can see that, the larger the $\beta_0$, the more power there is to detect the overall treatment effect. The parameter $\beta_1$ determines the treatment effect variability (increasing if $|\beta_1|$ gets larger). Larger $|\beta_1|$ implies that it is more likely to detect the subgroup. $\beta_0$ and $\beta_1$ together determine the overall treatment effect. The function `generate_y` can be used to generate the outcome values $y$.

To generate the binary treatment variable, we consider $$P(trt = 1|X) = \frac{1}{1+\exp(-h(X))}$$ where $h(X)$ is a user specified form. When $h(X) = 0$, $P(trt = 1|X) = 0.5$ represents a setting for complete randomized design (have a look at the function `generate_trt` for details).

Depending on different types of responses, the link function $g(\cdot)$ and how the clinical endpoints are generated are given as: 

- Continuous response: $g(x) = x$ and $Y \sim N(\mu = g(f(X)), \sigma)$.
- Binary response: $g(x) = \dfrac{e^x}{1 + e^x}$ and $Y \sim \text{Bernoulli}(p = g(f(X)))$.
- Count response: $g(x) = e^x$ and $Y\sim \text{Poisson}(\lambda = g(f(X)))$.
- Time-to-event response: $g(x) = e^x$ and $t\sim \text{Exponential}(\lambda = g(f(X)))$.

## Example of using ``benchtm'' package
In this section we provide an example of using "benchtm" package to generate a dataset for subgroup identification problem.

For the covariates, one can either utilize a user-specified dataset or generate them from the function `generate_X_dist` (generate covariates from a pre-specified distribution) or `generate_X_syn` (generate data from synthetic data from a specific real trial). See each corresponding function for more details.
```{r}
set.seed(1111)
X <- generate_X_dist(n=1000, p=10, rho=0.5)
#observational data set
trt <- generate_trt(n=nrow(X), p_trt = 0.5)
dat <- generate_y(X, trt, prog = "0.5*((X1=='Y')+X3)",
                   pred = "X3>0", b0 = 0, b1 = 1,
                   type = "continuous", sigma_error = 3)

head(dat)
```

The treatment is generated based on randomized trial where $P(trt = 1) = 0.5$. We could fit a conditional tree using this generated data. 
```{r}
cov.names <- dat %>% dplyr::select(starts_with("X")) %>% colnames()
eqn <- paste0("Y ~ trt| ",
              paste0(cov.names, collapse = "+"))

glmtr <- partykit::glmtree(as.formula(eqn), data = dat, family = "gaussian",
                           minsize = 0.2*nrow(dat))
plot(glmtr)
```

Based on the results, one can see that $X_3$ is selected as the split variable. The subgroup with $X_3>0.061$  has the largest treatment effect. 
