---
title: "Description of included synthetic data"
vignette: >
  %\VignetteIndexEntry{Description of included synthetic data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
# library(benchtm)
here::i_am("./vignettes/a04_synthetic_data.Rmd")
library(dplyr)
library(tidyverse)
```

As described in [Sun et al. (2024)](https://doi.org/10.1002/bimj.202100337) the data set is a synthetic data generated from a pool of clinical trial data. The ''synthpop'' function from the "synthpop" R package with default options was used to generate a synthetic version of the original data. For the sake of confidentiality we renamed the covariates into $X_1, X_2, \ldots, X_{30}$. In addition the factor levels of the categorical variables were renamed and the numeric variables were scaled into the interval $[0,1]$ by subtracting the smallest value and dividing by the range of values.  

In what follows we provide marginal summaries of the variables $X_1, X_2, ... X_{30}$ using barplots and density plots. In addition an assessment of the correlations among the numeric variables is provided.

```{r, warning = FALSE, echo = FALSE, message = FALSE, fig.show="hold", out.width="100%"}
load(here::here("./R/sysdata.rda"))
### function for density plot
library(ggplot2)
plot_density <- function(x.var, data){
  data %>% dplyr::select(x.var = x.var) %>%
    ggplot(aes(x = x.var)) + 
    geom_density(alpha = 0.25, color = "blue") + geom_rug() +
    theme_bw() + ylab("")+ 
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
                 legend.position = "none") + 
    ggtitle(x.var) + xlab("")
}
######### univariate bar plot
plot_bar <- function(x.var, data){
  ## x.var categorical, y.car: binary, bar plot based on y.var
   data %>% dplyr::select(x.var = x.var) %>%
    ggplot(aes(x.var)) +
   geom_bar(aes(y = after_stat(count/sum(count)), fill = "blue", alpha = 0.5)) +
    theme_bw() + theme(axis.text.x = element_text(angle = 30, hjust = 1),
                 legend.position = "none") +
    ylab("Percentage") + xlab("") +
    ggtitle(x.var)
}

pp_plots <- lapply(1:ncol(data_syn), function(ii){
  if(class(data_syn[[ii]]) == "numeric"){
    plot_density(colnames(data_syn)[ii], data_syn)
  }else{
    plot_bar(colnames(data_syn)[ii], data_syn)
  }
})

res <- gridExtra::marrangeGrob(pp_plots, nrow = 2, ncol = 4, layout_matrix = matrix(1:8, nrow = 2, byrow = T))
plot(res[[1]])
plot(res[[2]])
plot(res[[3]])
plot(res[[4]])
```

The Kendall's correlation between the continuous covariates $X_5,X_{10}$ - $X_{30}$ is shown in the correlation plot below.  

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.cap="Correlation for continuous variables for synthetic data"}
### continuous correlation
library(PerformanceAnalytics)
chart.Correlation(as.matrix(data_syn %>% dplyr::select(X5,X10:X30)), histogram=FALSE, pch=19, method = "kendall")
```

